Thesis: draft

### 2021-03-06 draft.tex
wrote section activation mechanisms + added some commentary notes + added one reference

\section{The activation mechanism}\label{actMech}

In the most simple opinion dynamics models, which makes use of a static network, users are always active and thus always exchanging opinions/posts with each other. In this kind of model the opinion dynamics might go as follows: each time step a random user is selected after which his/her opinion is updated according to the opinion dynamics model that is implemented. \\
It might, however, be a more realistic situation if nodes/users become active/inactive over time. This would be a better representation of real social media platforms, where user are active for some periods of time and then become inactive again. This kind of activation mechanism of the user gives the network a more temporal character. The underlying network structure doesn't change, but by making users active and inactive the exchange of information between users becomes time dependent and depends on which users are active and which aren't.\\
\newline
In this thesis we follow the same activation mechanism as was proposed in the paper of prof. Nicola Perra and prof. Luis E.C. Rocha (see reference \cite{Perra2019}). Every time step users are made active with some probability $p_i$ (thus, every time step, each node becomes active with a probability $p_i$). In this thesis the activation probability is set equal to $p = 0.1$, which is constant for all the nodes. Prof. Nicola Perra and prof. Luis E.C. Rocha choose, in their paper, activation probabilities that are extracted from a power-law ($F(p) \sim p^{-1.5}$, with $p \in \interval{0.01}{1}$) \cite{Perra2019}. They did this based on the analyses of real social networks that shows that the propensity of users to start an on-line social interaction follows a heterogeneous distribution \cite{Perra2019}. However, they also showed that using a constant activation probability $p$ over the entire population yields results that are qualitatively similar (at least in their case) \cite{Perra2019}. \\
Hence, for simplicity, this thesis implements the constant activation probability variant. It might however be an interesting extension to switch to the heterogeneous activation probabilities and investigate the possible impacts on the dynamics.

% not finished, read this https://www.nature.com/articles/srep00469 and https://www.nature.com/articles/srep03006, might contain useful info

@@

>>>>

\section{The activation mechanism: Static to temporal}\label{actMech}

In the most simple opinion dynamics models, which make use of a static network, users are always active and thus always exchanging opinions/posts with each other. In this kind of model the opinion dynamics might go as follows: each time step a random user is selected after which his/her opinion is updated according to the opinion dynamics model that is implemented. Another possibility is updating all users at the same time during each time step. \\
It might, however, be a more realistic situation if nodes/users become active/inactive over time. This would be a better representation of real social media platforms, where users are active for some periods of time and then become inactive again. This kind of activation mechanism of the users gives the network a more temporal character. The underlying network structure doesn't change (in the sense that each user/node has a fixed number of neighbors that doesn't change over time, so no new friends are made or, conversely, no old friends are being defriended), but by making users active and inactive the exchange of information between users becomes time dependent and depends on which users are active and which aren't. Hence, in this sense, this does affect the network structure in some way, because if a user, for example, doesn't become active for a long period of time, he/she doesn't participate in the exchange of information/opinions and the network structure behaves like if this user/node is absent. This may affect the dynamics, since it may seem as if the network has a different average degree and different things like connectivity etc. \\
In this thesis we follow the same activation mechanism as was proposed in the paper of prof. Nicola Perra and prof. Luis E.C. Rocha (see reference \cite{Perra2019}). Every time step users are made active with some probability $p_i$ (thus, every time step, each node becomes active with a probability $p_i$). In this thesis the activation probability is set equal to $p = 0.1$, which is constant for all the nodes. Prof. Nicola Perra and prof. Luis E.C. Rocha choose, in their paper, activation probabilities that are extracted from a power-law ($F(p) \sim p^{-1.5}$, with $p \in \interval{0.01}{1}$) \cite{Perra2019}. They did this based on the analyses of real social networks that shows that the propensity of users to start an on-line social interaction follows a heterogeneous distribution \cite{Perra2012}\cite{Perra2019}. However, they also showed that using a constant activation probability $p$ over the entire population yields results that are qualitatively similar (at least in their case) \cite{Perra2019}. \\
Hence, for simplicity, this thesis implements the constant activation probability variant. It might however be an interesting extension to switch to the heterogeneous activation probabilities and investigate the possible impacts on the dynamics. \\
If a user is active, he/she first spreads his current opinion to all his/her neighbors (for example by posting something). Then, the active user updates his own opinion according to the opinions on his/her time line \cite{Perra2019}. This updating goes according to the implemented opinion dynamics model (eg. majority model, probabilistic majority model,...). In every time step the updating of all active nodes is done simultaneously \cite{Perra2019}.\\
It is important to note that what somebody sees on his/her time line is the consequence of the interplay between the filtering algorithm and the posted opinions of his/her neighbors. Every user has a "hidden time line $L_i(t)$" which contains the actual time line of the user at the moment when he/she was active for the last time ($t_{\text{last}}$) plus all the posts of the neighbors from that last time he/she was active till the current time he/she becomes active again ($t$) \cite{Perra2019}. At the time $t$ that the user becomes active again, the filtering algorithm selects $Q$ post from this "hidden time line $L_i(t)$" and shows them on the actual time line of the user; these $Q$ items are thus the opinions that the user is exposed to \cite{Perra2019} (see also Section \ref{filter} for more detailed information on the filtering algorithms).

% NOT FOR HERE, but maybe in section over future recommendations, ways forward: talk about need to investigate different activation mechanisms, maybe model where nodes can make new friends and remove old friends (at a time rate comparable to the time rate of the opinion dynamics and also maybe based on the opinions of the friends, eg. if a friend has an opinion far from your, you might be more likely to defriend that one and, vice versa, you might be more likely to become new friends with somebody who has an opinion close to yours)

@@

% NOTE: maybe make somewhere in thesis link to real-world examples; for example the election system in USA: formation of echo chambers, different states voting for a different party
